Universal Anonymous Signatures

### 2022-09-19 prism-backend/research/gsac/formal-uas.tex
Completed in-depth pass. Some changes pending.

The main contribution of this section is giving a simple syntax and model that
still allow spanning a wide variety of related schemes. We combine aspects of GS
and AC schemes, generalizing whenever possible. Like in ACs, there are multiple
issuers, and each user can get one or more credentials from any of them; and,
also, use more than one credential simultaneously to produce a signature. We
incorporate openers, as in GSs, but there is no tight relationship between
issuer and opener. As a consequence, users (potentially along with verifiers)
may choose, on a per-signature basis, the policies governing the utility
information extractable from a signature. Finally, we incorporate functions that
allow tweaking the conditions checked at credential issuance and signature
generation times, including the information that an opener will be able to
extract from a signature.

\paragraph{Users and credentials.} %
In our model, every single user may have multiple credentials. In this
setting, a perhaps too strict assumption we make is that, once a user gets
compromised, all of the credentials owned by that user are compromised too.
While there may be real world situations for which this is too aggressive, it is
certainly a worst-case scenario, and greatly simplifies the modelling.
%
% Somehow related, we model a scheme in which users generate their secret keys in
% a standalone way, and then leverage these keys to request credentials. This may
% seem too ideal as a model for real world situations, as we cannot know how
% many keys a user has generated. However, we argue that this can in fact be very
% close to reality, where a well-known issuer, after properly identifying a user,
% attests the (possibly blinded) public key this user proves control of. Indeed,
% our model and syntax support exactly this approach, by allowing users to
% leverage previously obtained credentials in order to request new ones.
% A consequence of this is that, if ``one user'' generates (e.g.) two keys, and
% then uses each of these keys to request multiple credentials, then from the
% point of view of our modelling, we see this \emph{cryptographically} as two
% different users -- and, anyway, that same ``user'' cannot combine credentials
% bound to different keys.

\paragraph{Users and issuers.} %
A relevant aspect of our model is that users and issuers are basically the
same: an entity with a key pair. The difference is that, when users wish to
become issuers, they advertise their public keys, along with their conditions
for granting credentials. To simplify, we just assume that they somehow
``publish their public keys'', which can be done in several ways in practice.

\paragraph{Functional placeholders.} %
We use generic functions to capture, at the design level, that users need
to prove concrete claims over their private information. Looking a bit ahead,
the output of these functions is defined in terms of (subsets of) the users'
key, credentials and attributes, and signed message. But, during our modelling,
we are only concerned with preventing that the output of these functions can
be altered, and with ensuring that it does not threaten privacy beyond what is
necessary or unavoidable to reach the flexibility we want.

@@

An \UAS scheme is composed by the following algorithms:

\begin{description}
\item[$\Setup(\secpar) \rightarrow \parm$.] Given a security parameter \secpar,
  returns a global system parameter variable \parm. We assume that \parm are
  available to all other functions.
\item[$\KeyGen(\parm) \rightarrow (\upk,\usk)$.] Given \parm, a user runs
  \KeyGen to generate its user key pair. 
\item[$\OKeyGen(\parm,\finsp) \rightarrow (\opk,\osk)$.] Given \parm, and the
  function \finsp, an opener runs \OKeyGen to generate its opening key pair.
  The function \finsp defines the type of utility that will be extractable from
  signatures.
\item[$\ISet(\upk,\usk,\fissue) \rightarrow \ipk$.] Given a user key
  pair, and an issuance function \fissue for checking that credential requestors
  meet some conditions to be issued a credential, an existing user runs \ISet to
  extend the user public key, enabling this user to issue credentials meeting
  \fissue.
%\item[$\UKeyGen(\parm) \rightarrow (\upk,\usk)$.] Given \parm, returns a user's
%  key pair.
\item[$\langle
  \Obtain(\usk,\cidu,\ipk,\sCred,\attrs),
  \Issue(\isk,\cidi,\sipk,\attrs))  
  \rangle \rightarrow \langle \Cred/\bot,R/\bot
  \rangle$.] %
  This interactive protocol lets a user with key \usk receive a credential
  $\Cred=(\cid,\attrs,\cred,\ipk)$
  from an issuer in the system. $\cid=(\cidi,\cidu)$ is a unique identifier for
  the credential, where issuer and user randomly pick \cidi and \cidu,
  respectively (to ease notation, whenever we write just \cid, we mean $(\cidi,
  \cidu)$). \ipk is the issuer's public key and \cred is the actual credential,
  on attribute set \attrs. The user may leverage a set of previously obtained
  credentials $\sCred=\lbrace (\cid_i=(\cidi_i,\cidu_i),\cattrs_i,\cred_i,
  \ipk_i)\rbrace_{i\in[n]}$, from which we may omit the $\ipk_i$s for
  readability. The user outputs the produced credential \Cred, while the issuer
  outputs $R \gets (\yissue,\utrans)$, for some value $\yissue$ output by
  the \fissue function defined by the issuer, and $\utrans$ being the transcript
  of the communication.
\item[$\Sign(\usk,\opk,\sCred,\msg,\feval) \rightarrow (\sig,\yeval)$.] %
  Upon receiving user secret key \usk, opener public key \opk, 
  credentials \sCred, message \msg and evaluation function \feval, returns
  signature \sig, and a value \yeval. We use use \Sig to denote the
  tuple $(\sig,\yeval)$.
\item[$\Verify(\opk,\sipk,\Sig,\msg,\feval) \rightarrow 1/0$.]
  Checks whether $\Sig = (\sig,\yeval)$ is a valid signature
  over message \msg, from a user with credentials issued by issuers with public
  keys in \sipk, for evaluation function \feval and opener key \opk.
\item[$\Open(\osk,\sipk,\Sig,\msg,\feval) \rightarrow
  (\yinsp,\iproof)/\bot$.]
  Run by the opener with private key \osk. Receives a signature $\Sig=
  (\sig,\yeval)$ over message \msg and evaluation function \feval,
  generated using credentials by issuers with public keys in \sipk.
  If \Sig is valid, the function outputs a value $\yinsp$, and a proof of
  correct opening \iproof.
\item[$\Judge(\opk,\sipk,\yinsp,\iproof,\Sig,\msg,\feval) \rightarrow 1/0$.] %
  Checks if \iproof is a valid opening correctness proof for the value \yinsp,
  obtained by applying \Open to the the signature $\Sig = (\sig,\yeval)$
  over message \msg, and for evaluation function \feval. 
\end{description}

\paragraph{Issuance, evaluation, and opening functions.} %
We introduce three functions in our syntax and model. Through them, it is
possible to modulate the behaviour of the resulting instantiation, by
controling the conditions upon which a new credential can be issued, the
information revealed at signing time, or after signing. In all cases, it is
the user who computes them, and has to prove correctness of such computation.
We give concrete examples in \secref{sec:relationships}.

\begin{description}
\item[$\fissue: (\upk,\cid,\attrs,\lbrace (\cid_i,\cattrs_i)\rbrace_{i \in [n]})
  \rightarrow \yissue$.] Chosen
  by each issuer within a family of functions \famfissue, the issuance function
  defines the conditions required by the issuer to grant a credential
  with identifier $\cid = (\cidi,\cidu)$, over attributes \attrs, when requested
  by a user with public key \upk, leveraging a set of $n$ previously obtained
  credentials with identifiers and attributes given by $\lbrace (\cid_i,
  \cattrs_i) \rbrace_{i \in [n]}$ (where $n$ may be $0$). \fissue returns
  a value $\yissue \in \rngfissue \cup \lbrace 0 \rbrace$, where we adopt the
  convention that $\yissue = 0$ means that the issuance request is denied; and
  $\yissue \neq 0$ means that the request is approved. \fissue functions must be
  \UAS-acceptable, as defined in \defref{def:uas-acc-func}.
\item[$\feval: (\upk,\lbrace(\cid_i,\cattrs_i)\rbrace_{i \in [n]},\msg)
  \rightarrow (\yeval^0,\yeval^1)$.] Signing evaluation functions, from a
  family of functions \famfeval, can be set on a per-signature basis. They
  receive the user public key \upk, a set of credentials with identifiers
  and attributes $\lbrace (\cid_i,\cattrs_i) \rbrace_{i \in [n]}$
  (where $n \ge 1$), and the message to be signed \msg. \feval outputs two
  values, $\yeval^0$ and $\yeval^1$. Both must belong in a well defined set
  \rngfeval. For readability, we write $\Yeval = (\yeval^0,\yeval^1)$.
\item[$\finsp: (\Yeval,\upk,\lbrace (\cid_i,\cattrs_i) \rbrace_{i \in [n]},
  \msg) \rightarrow \yinsp$.]
  Chosen by openers from a family of functions \famfinsp. The opening
  functions define what utility value (derived from the user's secret key,
  credentials' identifiers and attributes used for signing, and signed message)
  should be extractable by an opener. Its behaviour can be modulated by \Yeval,
  and it outputs a value \yinsp, which must belong in a well defined set
  \rngfinsp.
\end{description}

\feval and \finsp may seem redundant. However, the former is chosen on a
per-signature basis, probably after negotiation between signer and verifier, and
governs the information revealed at signing time. On the other hand, \finsp
governs the information that may be revealed after signing, and is defined by
openers when they set up their key pair. Thus even though signers and verifiers
may negotiate to use any opener for a given signature, it is not possible for
an opener to change the chosen function.

@@

Next, we define the concept of \CUASGen restrictions. These are special
cases of our generic construction that target a concrete variant of the
utility-privacy tradeoff by defining concrete \fissue, \feval and \finsp
functions. Security of \CUASGen implies security of its restrictions, if
the chosen \fissue functions are \CUASGen-acceptable.

@@

We define the security expected from an \UAS scheme using a game-based approach.
An \UAS scheme must ensure privacy at credential issuance, meaning that two
runs of the credential issuance protocol should not leak any information beyond
necessary about the user requesting the credential. Note that this is now
crucial for privacy since \UAS allows multiple credentials per
user -- thus, knowing what user is requesting which credential can lead to user
profiling. It must also ensure signature privacy, meaning that no information
can be learned by third parties from a signature, beyond what the user chooses
to reveal, and what may be extractable by the chosen opener. Unforgeability
requires that \UAS schemes ensure that the claims used to support a credential
request cannot be falsified, and that no adversary is able to falsify
the correctness proofs for the claims included in the signatures. In addition,
we require all credentials involved in the corresponding process to be
associated to a known user key. Finally, \UAS schemes inherit the useful
non-frameability property of GS schemes where even the issuer is corrupt. In
this case, we can ensure that no honest user can be framed with having created a
signature that s/he did not create -- although the concept may get a bit
blurred, as there is no traditional opening.
%
We describe next the global variables and oracles that we use in our modelling.

\paragraph{Global Variables.} %
Users are referred to with user identifiers, \uid; for credentials, we use \cid;
and for openers, \oid. Although issuers are ``extended'' users, and thus can be
identified by their \uid, we also sometimes use \iid to refer to a user acting
as issuer, especially, for operations involving both a user and an issuer (this
also applies for key pairs).
%
For notational convenience, we often use these identifiers instead of the
corresponding user/issuer/opener keys.
%
All tables/sets are initialized as empty tables/sets. We use the following sets
and tables in our definitions:

\begin{description}
\item[$\mathsf{H}\lbrace\mathsf{U},\mathsf{I},\mathsf{O} \rbrace$.] Sets of
  honest {\uid}s/{\iid}s/{\oid}s.
\item[$\mathsf{C}\lbrace\mathsf{U},\mathsf{I},\mathsf{O} \rbrace$.] Sets of
  corrupt {\uid}s/{\iid}s/{\oid}s.
\item[\UK,\IK,\OK.] Tables maintaining key pairs of users $(\upk,\usk)$, issuers
  $(\ipk,\isk)$, and openers $(\opk,\osk)$. They are indexable by \uid,
  \iid, and \oid, respectively. Also, issuers' public keys contain their \fissue
  function, and openers' public keys contain their \finsp function.
\item[\PUBUK,\PUBIK,\PUBOK and \PRVUK,\PRVIK,\PRVOK.] Shorthands referring to
  public keys of users, issuers or openers ($\mathsf{PXK}$) or their private
  keys ($\mathsf{SXK}$).
\item[\CRED.] Stores information related to credentials obtained by users in
  the system. Thus, it is indexable by \cid. More specifically, it stores
  tuples of the form $(\uid,\cred,\iid,\attrs,n,\lbrace (\cid_i,\iid_i)
  \rbrace_{i \in [n]})$, where \uid is the identifier of the owner of the
  credential, \cred (when available) is the credential itself, \iid is the
  identifier of the credential issuer, \attrs are the attributes included in
  \cred, $n$ is the number of credentials used to support this request, and
  $(\cid_i,\iid_i)$ for $i \in [n]$ are the corresponding credential and issuer
  identifiers. For notational convenience, we may use $\CRED[\scid]$
  to refer to $\CRED[\cid]$ for all {\cid}s in a set \scid. Also, when clear
  from context, we sometimes use $\CRED[\cid]$ (resp. $\CRED[\scid]$) to
  mean \cred (resp. \scred) in $\CRED[\cid] = (\cdot,\cred,\cdot,\cdot,
  \cdot,\cdot)$ (resp. $\CRED[\scid]$).
\item[\CCRED.] Like \CRED, but tracks challenge credentials.
\item[\OWNR.] For notational convenience, when we write $\OWNR[\cid]$ we mean
  ``\uid such that $\CRED[\cid] = (\uid,\cdot,\cdot,\cdot,\cdot,\cdot)$'', or
  $\bot$ if there is no wuch \uid. We may also use \OWNR over a set of {\cid}s
  (\scid). In that case, our convention is that $\OWNR[\scid]$ returns $\bot$ if
  not all passed {\cid}s belong to the same \uid. Since $\bot$ is not a valid
  \uid, all conditions checking membership of the output of \OWNR to a list of
  users equal false if \OWNR returns $\bot$.
\item[\ATTR.] For notational convenience, when we write $\ATTR[\cid]$ we mean
  ``\attrs such that $\CRED[\cid] = (\cdot,\cdot,\cdot,\attrs,\cdot,\cdot)$''.
\item[\ISR.] For notational convenience, when we write $\ISR[\cid]$ we mean
  ``\iid such that $\CRED[\cid] = (\cdot,\cdot,\iid,\cdot,\cdot,\cdot)$''.
\item[\SIG.] Maintains signatures generated via the \SIGN oracle. Entries of
  this table are $(\oid,\scid,\sig,\Yeval,\msg,\feval)$ tuples, where \oid is
  the opener, \scid is the set of identifiers of the credentials used for
  signing, \feval is the signing evaluation function, and \sig and \msg are the
  produced signature and signed message.
\item[\CSIG.] Maintains challenge signatures output to the adversary in the
  signature anonymity game, and is indexable by challenge signatures \csig.
  Each entry contains also $\cuid_b$ and $\cscid_b$, respectively the challenge
  user  and credential identifiers set used to produce \csig; as well as the
  corresponding challenge user and credential set indexed by the complementary
  $1-b$; the signed message \msg and signing evaluation function \feval, the
  result of \feval, \Yeval, and the opener identifier \oid.
\end{description}

@@

\begin{description}
% \item[\IGEN.] Adds a new issuer to the game, generating its keypair and setting
%   the associated issuance function.
\item[\HUGEN.] Adds a new honest user to the game, generating its keypair.
\item[\CUGEN.] Adds a new corrupt user to the game, setting its \upk to the
  value given by the adversary.  
\item[\OGEN.] Adds a new opener to the game, generating its key pair and
  setting its evaluation and inspection functions.
% \item[\ICORR.] Corrupts an existing (and honest) issuer, by giving its secret
%   key to the adversary.
\item[\ISET.] ``Upgrades'' a user to become an issuer, specifying its
  \fissue function.
\item[\UCORR.] Corrupts an existing (and honest) user, giving its secret
  key to the adversary. If the user was also an issuer, it becomes a corrupt
  issuer too.
\item[\OCORR.] Corrupts an existing (and honest) opener, by giving its secret
  key to the adversary.
% \item[\HUGEN.] Adds a new honest user to the game, by honestly generating
%   the user's key pair.
% \item[\CUGEN.] Adds a new corrupt user to the game or, if the specified
%   user already exists and is honest, corrupts it, leaking its key and
%   credentials.
\item[\RREG.] Reads the given transcript table entry.
\item[\WREG.] Sets a transcript table entry to the given value.
\item[\OBTISS.] Lets the adversary add a new honestly generated credential to
  the game, on behalf of an honest user.
\item[\OBTAIN.] Enables the adversary to play the role of a dishonest issuer, by
  letting it interact with honest users who want to receive credentials.
\item[\ISSUE.] Allows the adversary to play the role of dishonest users,
  requesting an honest issuer to produce credentials for them.
\item[\SIGN.] Lets the adversary get signatures using credentials from honest
  users.
\item[\OPEN.] Given an honestly produced signature, outputs the result of the
  opening function, along with a correctness proof.
\item[\OBTCHALb.] Upon receiving a target credential identifier, two challenge
  users and credential sets, an honest issuer identifier, and a set of
  attributes, issues a credential for the challenge user and credential set
  defined by the bit $b$. 
\item[\CHALb.] Upon receiving two challenge users and credential sets, a common
  singing evaluation function and a message, returns a signature produced by one
  of these two user and credential sets, defined by the bit $b$.
\end{description}

@@

        \procedure{$\OBTISS(\stackrel{\cid}{\pcbox{(\cidi,\cidu)}},\uid,\iid,\attrs,
          \lbrace (\cid_i,\iid_i) \rbrace_{i \in [n]})$}{%
          \pcif \uid \notin \HU \lor \iid \notin \HI: \pcreturn \bot \\
          \pcif \exists i \in [n]~\suchthat~\cid_i \in \CCRED: \pcreturn \bot \\
          \pcif \CRED[\cid] \neq \bot \lor \CCRED[\cid] \neq \bot:
          \pcreturn \bot \\
          \langle \Cred, R \rangle \gets
          \langle \Obtain(\UK[\uid],\cidu,\CRED[\lbrace \cid_i\rbrace_{i\in[n]}],\attrs), \\
          \hspace*{50pt} \Issue(\PRVIK[\iid],\cidi,\ISR[\lbrace \cid_i\rbrace_{i\in[n]}],\attrs)
          \rangle \\
          \trans[\cid] \gets R \\
          \CRED[\cid] \gets (\uid, \cred, \iid, \attrs, n,
          \lbrace(\cid_i,\iid_i)\rbrace_{i\in[n]}) \\
          \pcreturn \top \\
        }

>>>>

The main contribution of this section is giving a simple syntax and model that
still allow spanning a wide variety of related schemes. This is in part
achieved by combining aspects of GS and AC schemes. Yet, the main tool for
generalisation are three functional placeholders, governing the conditions to be
checked during issuance, the information to be revealed alongside signatures,
and extractable from them. However, at the modelling level we are not concerned
with the ``code'' of these functions, but only with the correctness of their
computation, and with capturing security and privacy despite introducing
configurable functions.

% \paragraph{Users and credentials.} %
% In our model, every single user may have multiple credentials. In this
% setting, a perhaps too strict assumption we make is that, once a user gets
% compromised, all of the credentials owned by that user are compromised too.
% While there may be real world situations for which this is too aggressive, it is
% certainly a worst-case scenario, and greatly simplifies the modelling.
% %
% % Somehow related, we model a scheme in which users generate their secret keys in
% % a standalone way, and then leverage these keys to request credentials. This may
% % seem too ideal as a model for real world situations, as we cannot know how
% % many keys a user has generated. However, we argue that this can in fact be very
% % close to reality, where a well-known issuer, after properly identifying a user,
% % attests the (possibly blinded) public key this user proves control of. Indeed,
% % our model and syntax support exactly this approach, by allowing users to
% % leverage previously obtained credentials in order to request new ones.
% % A consequence of this is that, if ``one user'' generates (e.g.) two keys, and
% % then uses each of these keys to request multiple credentials, then from the
% % point of view of our modelling, we see this \emph{cryptographically} as two
% % different users -- and, anyway, that same ``user'' cannot combine credentials
% % bound to different keys.

% \paragraph{Users and issuers.} %
% A relevant aspect of our model is that users and issuers are basically the
% same: an entity with a key pair. The difference is that, when users wish to
% become issuers, they advertise their public keys, along with their conditions
% for granting credentials. To simplify, we just assume that they somehow
% ``publish their public keys'', which can be done in several ways in practice.

@@

Users generate their key pair, and then optionally advertise
their public key and conditions for issuance in order to become issuers. When
openers generate their key pairs, they include the function they will apply to
anonymous signatures. Users can later leverage (one or more of) their
credentials to request new ones, or to produce an anonymous signature. Issuance
only succeeds if the user's credentials (if any) meet the issuer's requirements.
Signatures may directly output some information derived from the user's data.
Also, each signature has a selected opener (probably, by the user and/or
verifier), who can later \emph{extract only the information that is defined by
  the chosen opener's key}. When openers extract their information, they also
prove the correctness of the result.

In detail, an \UAS scheme is composed by the following algorithms:

\begin{description}
\item[$\Setup(\secpar) \rightarrow \parm$.] Given a security parameter \secpar,
  returns a global system parameter variable \parm. We assume that \parm are
  available to all other functions.
\item[$\KeyGen(\parm) \rightarrow (\upk,\usk)$.] Given \parm, a user runs
  \KeyGen to generate its user key pair. 
\item[$\OKeyGen(\parm,\finsp) \rightarrow (\opk,\osk)$.] Given \parm, and the
  function \finsp, an opener runs \OKeyGen to generate its opening key pair.
  The function \finsp defines the type of utility that will be extractable from
  signatures.
\item[$\ISet(\upk,\usk,\fissue) \rightarrow \ipk$.] Given a user key
  pair, and an issuance function \fissue for checking that credential requestors
  meet some conditions to be issued a credential, an existing user runs \ISet to
  extend the user public key, enabling this user to issue credentials meeting
  \fissue.
%\item[$\UKeyGen(\parm) \rightarrow (\upk,\usk)$.] Given \parm, returns a user's
%  key pair.
\item[$\langle
  \Obtain(\usk,\ipk,\sCred,\attrs),
  \Issue(\isk,\sipk,\attrs))  
  \rangle \rightarrow \langle \Cred/\bot,R/\bot
  \rangle$.] %
  This interactive protocol lets a user with key \usk receive a credential
  $\Cred=(\cid,\attrs,\cred,\ipk)$
  from an issuer in the system. $\cid=(\cidi,\cidu)$ is a unique identifier for
  the credential, where issuer and user randomly pick \cidi and \cidu,
  respectively (to ease notation, whenever we write just \cid, we mean $(\cidi,
  \cidu)$). \ipk is the issuer's public key and \cred is the actual credential,
  on attribute set \attrs. The user may leverage a set of previously obtained
  credentials $\sCred=\lbrace (\cid_i=(\cidi_i,\cidu_i),\cattrs_i,\cred_i,
  \ipk_i)\rbrace_{i\in[n]}$, from which we may omit the $\ipk_i$s for
  readability. The user outputs the produced credential \Cred, while the issuer
  outputs $R \gets (\yissue,\utrans)$, for some value $\yissue$ output by
  the \fissue function defined by the issuer, and $\utrans=(U,I)$ being the
  transcript of the communication, where $U$ contains the messages sent by
  the user, and $I$ those sent by the issuer.
\item[$\Sign(\usk,\opk,\sCred,\msg,\feval) \rightarrow (\sig,\yeval)$.] %
  Upon receiving user secret key \usk, opener public key \opk, 
  credentials \sCred, message \msg and evaluation function \feval, returns
  signature \sig, and a value \yeval. We use use \Sig to denote the
  tuple $(\sig,\yeval)$.
\item[$\Verify(\opk,\sipk,\Sig,\msg,\feval) \rightarrow 1/0$.]
  Checks whether $\Sig = (\sig,\yeval)$ is a valid signature
  over message \msg, from a user with credentials issued by issuers with public
  keys in \sipk, for evaluation function \feval and opener key \opk.
\item[$\Open(\osk,\sipk,\Sig,\msg,\feval) \rightarrow
  (\yinsp,\iproof)/\bot$.]
  Run by the opener with private key \osk. Receives a signature $\Sig=
  (\sig,\yeval)$ over message \msg and evaluation function \feval,
  generated using credentials by issuers with public keys in \sipk.
  If \Sig is valid, the function outputs a value $\yinsp$, and a proof of
  correct opening \iproof.
\item[$\Judge(\opk,\sipk,\yinsp,\iproof,\Sig,\msg,\feval) \rightarrow 1/0$.] %
  Checks if \iproof is a valid opening correctness proof for the value \yinsp,
  obtained by applying \Open to the the signature $\Sig = (\sig,\yeval)$
  over message \msg, and for evaluation function \feval. 
\end{description}

\paragraph{Issuance, evaluation, and opening functions.} %
These are the functional placeholders that allow modulating the behaviour of the
resulting instantiation, by controling the conditions upon which new credentials
can be issued, the information revealed at signing time, or extractable from
signatures. In all cases, the user who computes them, and proves correctness of
such computation. We give concrete examples in \secref{sec:relationships}.

\begin{description}
\item[$\fissue: (\upk,\cid,\attrs,\lbrace (\cid_i,\cattrs_i)\rbrace_{i \in [n]})
  \rightarrow \yissue$.] Chosen
  by each issuer within a family of functions \famfissue, the issuance function
  defines the conditions required by the issuer to grant a credential
  with identifier $\cid = (\cidi,\cidu)$, over attributes \attrs, when requested
  by a user with public key \upk, leveraging a set of $n$ previously obtained
  credentials with identifiers and attributes given by $\lbrace (\cid_i,
  \cattrs_i) \rbrace_{i \in [n]}$ (where $n$ may be $0$). \fissue returns
  a value $\yissue \in \rngfissue \cup \lbrace 0 \rbrace$, where we adopt the
  convention that $\yissue = 0$ means that the issuance request is denied; and
  $\yissue \neq 0$ means that the request is approved. \fissue functions must be
  \UAS-acceptable, as defined in \defref{def:uas-acc-func}.
\item[$\feval: (\upk,\lbrace(\cid_i,\cattrs_i)\rbrace_{i \in [n]},\msg)
  \rightarrow (\yeval^0,\yeval^1)$.] Signing evaluation functions (or, simply,
  evaluation functions), from a
  family of functions \famfeval, can be set on a per-signature basis. They
  receive the user public key \upk, a set of credentials with identifiers
  and attributes $\lbrace (\cid_i,\cattrs_i) \rbrace_{i \in [n]}$
  (where $n \ge 1$), and the message to be signed \msg. \feval outputs two
  values, $\yeval^0$ and $\yeval^1$. Both must belong in a well defined set
  \rngfeval. For readability, we write $\Yeval = (\yeval^0,\yeval^1)$.
\item[$\finsp: (\Yeval,\upk,\lbrace (\cid_i,\cattrs_i) \rbrace_{i \in [n]},
  \msg) \rightarrow \yinsp$.]
  Chosen by openers from a family of functions \famfinsp. The opening
  functions define what utility value (derived from the user's secret key,
  credentials' identifiers and attributes used for signing, and signed message)
  should be extractable by an opener. Its behaviour can be modulated by \Yeval,
  and it outputs a value \yinsp, which must belong in a well defined set
  \rngfinsp.
\end{description}

\feval and \finsp may seem redundant. However, the former is chosen on a
per-signature basis, probably after negotiation between signer and verifier, and
governs the information revealed at signing time. On the other hand, \finsp
governs the information that may be revealed after signing, and is defined by
openers when they set up their key pair.

@@

Next, we define the concept of \CUASGen restrictions. These are special
cases of our generic construction that target a concrete variant of the
utility-privacy tradeoff by defining concrete \fissue, \feval and \finsp
functions. Security of \CUASGen implies security of its restrictions, if
the chosen \fissue function is \CUASGen-acceptable.

@@

We define the security expected from an \UAS scheme using a game-based approach.
An \UAS scheme must ensure privacy at credential issuance, meaning that two
runs of the credential issuance protocol should not leak any information beyond
necessary about the user requesting the credential. Note that this is now
crucial for privacy since \UAS allows multiple credentials per
user -- thus, knowing what user is requesting which credential can lead to user
profiling. It must also ensure signature privacy, meaning that no information
can be learned by third parties from a signature, beyond what the user chooses
to reveal, and what may be extractable by the chosen opener. Unforgeability
requires that \UAS schemes ensure that the claims used to support a credential
request cannot be falsified, and that no adversary is able to falsify the
correctness proofs for the claims included in the signatures. In addition, we
require all credentials involved in the corresponding process to be associated
to the same, and known, user key. Finally, \UAS schemes inherit the useful
non-frameability property of GS schemes where even the issuer is corrupt. In
this case, we can ensure that no honest user can be framed with having created a
signature that s/he did not create -- although the concept may get a bit
blurred, as there is no traditional opening.
%
We describe next the global variables and oracles that we use in our modelling.

\paragraph{Global Variables.} %
Users are referred to with user identifiers, \uid; for credentials, we use \cid;
and for openers, \oid. Although issuers are ``extended'' users, and thus can be
identified by their \uid, we mostly use \iid to refer to a user acting as
issuer, especially, for operations involving both a user and an issuer.
%
For notational convenience, we often use these identifiers instead of the
corresponding user/issuer/opener keys.
%
All tables/sets are initialized as empty tables/sets. We use the following sets
and tables in our definitions:

\begin{description}
\item[$\mathsf{H}\lbrace\mathsf{U},\mathsf{I},\mathsf{O} \rbrace$.] Sets of
  honest {\uid}s/{\iid}s/{\oid}s.
\item[$\mathsf{C}\lbrace\mathsf{U},\mathsf{I},\mathsf{O} \rbrace$.] Sets of
  corrupt {\uid}s/{\iid}s/{\oid}s.
\item[\UK,\IK,\OK.] Tables storing key pairs of users $(\upk,\usk)$, issuers
  $(\ipk,\isk)$, and openers $(\opk,\osk)$, indexable by \uid, \iid, and \oid,
  respectively. Issuers' public keys contain a \fissue function, and
  openers' public keys contain a \finsp function.
\item[\PUBUK,\PUBIK,\PUBOK and \PRVUK,\PRVIK,\PRVOK.] Shorthands referring to
  public keys of users, issuers or openers ($\mathsf{PXK}$) or their private
  keys ($\mathsf{SXK}$).
\item[\CRED.] Indexable by \cid, it stores
  tuples of the form $(\uid,\cred,\iid,\attrs,n,\lbrace (\cid_i,\iid_i)
  \rbrace_{i \in [n]})$, where \uid is the identifier of the owner of the
  credential, \cred (when available) is the credential itself, \iid is the
  identifier of the credential issuer, \attrs are the attributes included in
  \cred, $n$ is the number of credentials used to support this request, and
  $(\cid_i,\iid_i)$ for $i \in [n]$ are the corresponding credential and issuer
  identifiers. For notational convenience, we may use $\CRED[\scid]$
  to refer to $\CRED[\cid]$ for all {\cid}s in a set \scid. Also, when clear
  from context, we sometimes use $\CRED[\cid]$ (resp. $\CRED[\scid]$) to
  mean \cred (resp. \scred) in $\CRED[\cid] = (\cdot,\cred,\cdot,\cdot,
  \cdot,\cdot)$ (resp. $\CRED[\scid]$).
\item[\CCRED.] Like \CRED, but tracks challenge credentials.
\item[\OWNR.] When we write $\OWNR[\cid]$ we mean
  ``\uid such that $\CRED[\cid] = (\uid,\cdot,\cdot,\cdot,\cdot,\cdot)$'', or
  $\bot$ if there is no wuch \uid. We also use \OWNR over a set of {\cid}s
  (\scid). In that case, $\OWNR[\scid]$ returns $\bot$ if
  not all passed {\cid}s belong to the same \uid. Since $\bot$ is not a valid
  \uid, all conditions checking membership of the output of \OWNR to a list of
  users equal false if \OWNR returns $\bot$.
\item[\ATTR.] When we write $\ATTR[\cid]$ we mean
  ``\attrs such that $\CRED[\cid] = (\cdot,\cdot,\cdot,\attrs,\cdot,\cdot)$''.
\item[\ISR.] When we write $\ISR[\cid]$ we mean
  ``\iid such that $\CRED[\cid] = (\cdot,\cdot,\iid,\cdot,\cdot,\cdot)$''.
\item[\SIG.] Maintains signatures generated via the \SIGN oracle. Entries of
  this table are $(\oid,\scid,\sig,\Yeval,\msg,\feval)$ tuples, where \oid is
  the opener, \scid is the set of identifiers of the credentials used for
  signing, \feval is the signing evaluation function, and \sig and \msg are the
  produced signature and signed message.
\item[\CSIG.] Maintains challenge signatures output to the adversary in the
  signature anonymity game, and is indexable by challenge signatures \csig.
  Each entry contains also $\cuid_b$ and $\cscid_b$, respectively the challenge
  user  and credential identifiers set used to produce \csig; as well as the
  corresponding challenge user and credential set indexed by the complementary
  $1-b$; the signed message \msg and signing evaluation function \feval, the
  result of \feval, \Yeval, and the opener identifier \oid.
\end{description}

@@

\begin{description}
% \item[\IGEN.] Adds a new issuer to the game, generating its keypair and setting
%   the associated issuance function.
\item[\HUGEN.] Adds a new honest user to the game, generating its keypair.
\item[\CUGEN.] Adds a new corrupt user to the game, setting its \upk to the
  value given by the adversary.  
\item[\OGEN.] Adds a new opener to the game, generating its key pair and
  setting its evaluation and opening functions.
% \item[\ICORR.] Corrupts an existing (and honest) issuer, by giving its secret
%   key to the adversary.
\item[\ISET.] ``Upgrades'' a user to become an issuer, specifying its
  \fissue function.
\item[\UCORR.] Corrupts an existing (and honest) user, giving its secret
  key to the adversary. If the user was also an issuer, it becomes a corrupt
  issuer too.
\item[\OCORR.] Corrupts an existing (and honest) opener, by giving its secret
  key to the adversary.
% \item[\HUGEN.] Adds a new honest user to the game, by honestly generating
%   the user's key pair.
% \item[\CUGEN.] Adds a new corrupt user to the game or, if the specified
%   user already exists and is honest, corrupts it, leaking its key and
%   credentials.
\item[\RREG.] Reads the given transcript table entry.
\item[\WREG.] Sets a transcript table entry to the given value.
\item[\OBTISS.] Lets the adversary add a new honestly generated credential to
  the game, on behalf of an honest user.
\item[\OBTAIN.] Enables the adversary to play the role of a dishonest issuer, by
  letting it interact with honest users who want to receive credentials.
\item[\ISSUE.] Allows the adversary to play the role of dishonest users,
  requesting an honest issuer to produce credentials for them.
\item[\SIGN.] Lets the adversary get signatures using credentials from honest
  users.
\item[\OPEN.] Given an honestly produced signature, outputs the result of the
  opening function, along with a correctness proof.
\item[\OBTCHALb.] Upon receiving a target credential identifier, two challenge
  users and credential sets, an honest issuer identifier, and a set of
  attributes, issues a credential for the challenge user and credential set
  defined by the bit $b$. 
\item[\CHALb.] Upon receiving two challenge users and credential sets, a common
  singing evaluation function and a message, returns a signature produced by one
  of these two user and credential sets, defined by the bit $b$.
\end{description}

@@

        \procedure{$\OBTISS(\uid,\iid,\attrs,\lbrace (\cid_i,\iid_i)
          \rbrace_{i \in [n]})$}{%
          \pcif \uid \notin \HU \lor \iid \notin \HI: \pcreturn \bot \\
          \pcif \exists i \in [n]~\suchthat~\cid_i \in \CCRED: \pcreturn \bot \\          
          \langle \Cred, R \rangle \gets
          \langle \Obtain(\UK[\uid]\CRED[\lbrace \cid_i\rbrace_{i\in[n]}],\attrs), \\
          \hspace*{50pt} \Issue(\PRVIK[\iid],\ISR[\lbrace \cid_i\rbrace_{i\in[n]}],\attrs)
          \rangle \\
          (\cid = (\cidi,\cidu),\cdot,\cdot,\cdot) \gets \Cred \\
          \pcif \CRED[\cid] \neq \bot \lor \CCRED[\cid] \neq \bot:
          \pcreturn \bot \\
          \trans[\cid] \gets R \\
          \CRED[\cid] \gets (\uid, \cred, \iid, \attrs, n,
          \lbrace(\cid_i,\iid_i)\rbrace_{i\in[n]}) \\
          \pcreturn \top \\
        }
